DeconvOptim.jl - Signal Deconvolution with Julia

### 2021-08-07 paper/paper.tex

\begin{abstract}
\todo{fix citations with PkgCite.jl}
Deconvolution is a versatile method to enhance the quality of signals measured with systems which can 
be expressed mathematically as a convolution of a system's response function with a signal. 
In this paper, we present \verb|DeconvOptim.jl|, a flexible toolbox written in Julia to deconvolve one or multiple multi-dimensional signals which have been
degraded by a multi-dimensional signal response function. 
\verb|DeconvOptim.jl| works both on CPUs and GPUs and utilizes recent advancements in Julias automatic differentiation ecosystem.

@@

\todo{multi view deconvolution}

@@

\subsection{Sliding Convolution Windows}
Convolutions can be also calculated with a sliding kernel in real space which is the direct approach to
solve the integral in \autoref{eq:conv}.
That mechanism is the working principle
behind the convolutions in neural nets. The reason is that this operation can be parallelized easily and therefore 
is suitable on GPUs. However, the computational complexity for a 2D kernel with size $K \times L$ applied on an image with
$N \times M$ data points is $\mathcal O(N \cdot M \cdot K \cdot L)$. It is clear that for larger kernel sizes ($K,L \gg 1$) the sliding method 
has worse computational complexity than the FFT based convolutions. The many cores of a GPU can mitigate that to a certain extent
but since the FFT (CuFFT in the CUDA library \cite{CUDA}) can be also executed in parallel fashion, both approaches profit from parallelization.
If the kernel is separable over the dimensions (like a Gaussian kernel) the complexity changes to $\mathcal O(N \cdot M \cdot L +  N \cdot M \cdot K)$ however most optical PSFs are not separable and therefore this is not generally applicable.

\subsection{Wrap-Around Artifacts in FFT based Convolutions}
    One drawback of a FFT based convolution are wrap-around artifacts (see arrow in \autoref{fig:withoutpad}). The reason is, that the FFT operation calculates the discrete
    Fourier transform (DFT) which inherently assumes periodic data. If one applies a FFT on a finite data set the assumption is that
    this finite data set repeats itself. 
    To weaken the wrap-around artifacts in the convolution one can zero or mean pad the image before convolution 
    and remove the padding afterwards. Such approaches still suffer from artifacts when boundary values do not fit
    to the mean or to zero as it can be observed in \autoref{fig:withpad} where the arrow is pointing to.
    \begin{figure}[h]
        \centering
        \begin{subfigure}[b]{.25\textwidth}
            \centering
            \begin{tikzpicture}
                \node at (0,0) {\includegraphics[width=.9\textwidth]{figures/full_house_no_pad.png}};
                \draw[red] (-2.0, -1.6) rectangle (-1.6, -1.2);
                \node[color=red] at (0,0) {\includegraphics[width=.4\textwidth]{figures/edge_house_no_pad.png}};
                \draw[color=red] (-0.91, -0.9) rectangle (0.9, 0.9); 
                \draw[color=red] (-2.0, -1.2) -- (-0.9, 0.9);
                \draw[color=red] (-1.6, -1.6) -- (0.9, -0.9);
                \draw[<-, color=red, thick ](-0.85, 0) -- (-0.4, 0.6);
            \end{tikzpicture}
            \caption{Without padding}
            \label{fig:withoutpad}
        \end{subfigure}%
        \begin{subfigure}[b]{.25\textwidth}
            \centering
            \begin{tikzpicture}
                \node at (0,0) {\includegraphics[width=.9\textwidth]{figures/full_house_pad.png}};
                \draw[red] (-2.0, -1.6) rectangle (-1.6, -1.2);
                \node[color=red] at (0,0) {\includegraphics[width=.4\textwidth]{figures/edge_house_pad.png}};
                \draw[color=red] (-0.90, -0.9) rectangle (0.9, 0.9); 
                \draw[color=red] (-2.0, -1.2) -- (-0.9, 0.9);
                \draw[color=red] (-1.6, -1.6) -- (0.9, -0.9);
                \draw[<-, color=red, thick ](-0.8, -0.7) -- (-0.3, -0.3);
            \end{tikzpicture}
            \caption{With padding.}
            \label{fig:withpad}
        \end{subfigure}
        \caption{Comparison between two blurred images. The image a) is a straightforward FFT based blurred image.
        The zoomed region shows dark artifacts at the left side originated from the right hand side. Also at the top of the image some dark regions from
        the bottom scattered into the bright sky.
        Image b) was padded before blurring with 10 pixels on each side with value 0.7 which results in bright regions around the image. The blur kernel was a Gaussian kernel
        with standard deviation $\sigma=\SI{3}{pixel}$.}
        \label{fig:}
    \end{figure}
    To solve for the wrap-around artifacts in deconvolution we use a different approach which will be presented in \autoref{sec:wrap}. 
\subsection{Preferred Convolution}
The current state of \verb|DeconvOptim.jl| uses the FFT approach to calculate the convolution.
However, in principle one could add the possibility to choose spatially based convolutions.
But since the runtime gets especially long in multi-dimensional signal deconvolution where FFT based methods
are faster, we always use FFT based convolutions here.

\section{Inverse Modelling}
In \autoref{fig:pipeline} we show our approach to deconvolution.
The general idea is to create a (physical) forward model describing the measurement process.
\begin{figure}[h]
    \centering
    \includegraphics[width = .5\textwidth]{figures/pipeline.pdf}
    \caption{}
    \label{fig:pipeline}
\end{figure}

At the beginning of the optimization we start with an initial guess $x(\mathbf r)$. 
Via a mapping function we can impose some restrictions to the final solution (e.g. non-negativity, value intervals).
The forward model is then applied to the mapped reconstruction. We provide a default function where 
the forward function is a convolution but we also allow to plug in a different forward model.
This output of the forward function is compared
to our noisy measurement under a given loss function.
Since deconvolution is an ill-posed problem, we regularize the loss function with a regularizer.
Afterwards we calculate the gradient of the loss function with respect to the reconstruction $x(\mathbf r)$ via automatic differentation (AD).
The values of the total loss and the gradient are then used by an optimizer (to minimize the total loss) to apply a gradient descent optimization to get a proper reconstruction.\\

As typical for Julia packages, except \verb|FFTW.jl| and \verb|CUDA.jl| \cite{besard2018juliagpu}, all code is written in Julia.
Therefore it is easy to replace the forward model with any Julia code, as long as we can 
automatically differentiate through the code. The gradient step is currently calculated with the reverse mode AD of \verb|Zygote.jl| \cite{Zygote.jl-2018}. 
As default for the optimization we use the L-BFGS \cite{LBFGS} routine provided by \verb|Optim.jl| \cite{mogensen2018optim}.
L-BFGS is suitable for optimization with many million parameters as it only stores a sparse representation of the inverse Hessian matrix,
which is built from a finite number of past function and gradient evaluations. 
Results are often good after 10 to 30 iterations whereas Richardson-Lucy takes
a few hundreds iterations. 
However, as we will see later due to the more complex optimization routine the speed-up in using L-BFGS is for a simple 
deconvolution minor.

\subsection{Mapping Function}
    Mapping functions are an easy way to constrain the resulting reconstruction to predefined ranges.
    For example, if the optimizer optimizes the variables $\text{rec}$ which are then squared, we know that
    $\text{rec}^2$ is definitely non-negative. Instead of directly using $\text{fwd}(\text{rec})$ we compare
    $\text{fwd}(\text{rec}^2)$ with the measurement $Y$. This mapping function does not change the outcome as long
    as it is an surjective function for the desired number space (e.g. a parabola is non-negative and covers all positive values) 
    and does not change the convexity of a problem.
    Functions like shifted $\tanh$ are possible as well since it is an bijective function in $[-1, 1]$ and monotonic increasing
    in the whole definition space.
    Hence the final result is $\text{rec}^2$ in case of the quadratic mapping.

\subsection{Wrap-around Artifacts Suppression}
    \label{sec:wrap}
    To prevent that wrap-around artifacts occur during deconvolution we can instead
    of reconstructing an object having the same size as the measurement, reconstruct
    an object having a slightly larger size. This is shown in \autoref{fig:wrap_around}.
    \begin{figure}[h]
        \centering
        \begin{tikzpicture}
            \draw[fill=red!20!white] (0,0) rectangle (4,4);
            \node[color=red] at (2, 3.8) {Padded area};
            \draw[fill=gray!40!white] (0.5,0.5) rectangle (3.5,3.5); 
            \node[text width=2.5cm, color=black] at (2, 2) {area compared with measurement};
        \end{tikzpicture}
        \caption{Reconstructed data is padded before deconvolution but only the red region is compared with the measurement.}
        \label{fig:wrap_around}
    \end{figure}
    Consequently, we only compare the grey part of the reconstruction with the measurement.
    Through the AD and the convolution operations we still optimize for the whole area (including the red one).
    With that padding we reduce that images on the boundary influence in circular manner on reconstruction in the grey area.

\section{Loss Function}
    Apart from forward model, the loss function is the most critical part in the optimization.
    It has to account for the noise created by the measurement process or device.

@@

\section{Experiments}
    In this section we show some experiments and compare parts of the software to other solutions.
    All experiments were executed on Ubuntu 20.04 using a AMD Ryzen 5 5600X 6-Core processor (12 threads) with \SI{32}{GB} DDR4 RAM and Julia 1.6.1.
    If possible, all threads were used.
    As GPU accelerator we used a GeForce RTX 2060 SUPER with \SI{8}{GB} memory and CUDA 11.3 and CUDA.jl 3.3.1.

    \subsection{One-Dimensional Time Series}
        First, we want to demonstrate the deconvolution of an artificial one dimensional dataset blurred with a certain PSF.
        Additionally, we want to demonstrate the mechanism to reduce wrap-around artifacts in the deconvolution.
        On the left hand side in \autoref{fig:1D_dataset} we can see a data series which has been simulated 
        as a convolution with a PSF and degradation with additive Gaussian noise.
        \begin{figure}[H]
            \begin{subfigure}[b]{.25\textwidth}
                \centering
                \includegraphics[width=\textwidth]{figures/1d_plot_data_blurred.pdf}
                \caption{}
                \label{fig:1D_dataseta}
            \end{subfigure}%
            \begin{subfigure}[b]{.25\textwidth}
                \centering
                \includegraphics[width=\textwidth]{figures/1d_plot_data_res_pad_vs_nopad.pdf}
                \caption{}
                \label{fig:1D_datasetb}
            \end{subfigure}
            \caption{1D data which has been convolved with a PSF and was degraded with additive Gaussian noise with $\sigma=1.0$. Note that for the deconvolution we have chosen a position interval which is well within the full series so that there are not wrap-around artifacts in the measurement.}
            \label{fig:1D_dataset}
        \end{figure}
        On the right hand we can see the results of the following deconvolution:
        \begin{lstlisting}[language=Julia]
r_p, o = deconvolution(measured, psf, loss=Gauss(),
     iterations=12, regularizer=TV(num_dims=1),
     λ=0.05, padding=1)
r, o = deconvolution(measured, psf, loss=Gauss(), 
     iterations=12, regularizer=TV(num_dims=1),
     λ=0.05, padding=0)
        \end{lstlisting}
    The first statement applies a total padding of $1$ meaning that the size is doubled and therefore the measured data is padded.
    As we can see, the blue curve in \autoref{fig:1D_datasetb} does not show these intense spikes at the boundaries as the green curve shows.
    Also we notice that the deconvolved data in \autoref{fig:1D_datasetb} does not show all the details which are visible in the ground truth.
    The reason is that the PSF is a low pass filter and therefore only frequencies which have been passed through the filter 
    can be enhanced in terms of contrast in a deconvolved result.

    \subsection{Influence of Regularizer}
        In this part we want to demonstrate the influence of the regularizer.
        \autoref{fig:fabioa} is a simulated imaged which has been affected by Poisson shot noise and blurring by a PSF.
        \begin{figure}[h]
            \begin{subfigure}[b]{.25\textwidth}
                \centering
                \includegraphics[width=\textwidth]{figures/resolution_test_chart_blurry.png}
                \caption{Measured with Poisson noise}
                \label{fig:fabioa}
            \end{subfigure}%
            \hfill
            \begin{subfigure}[b]{.25\textwidth}
                \centering
                \includegraphics[width=\textwidth]{figures/resolution_test_chart_no_reg_iter_50.png}
                \caption{Without regularizer $N=50$}
                \label{fig:fabiob}
            \end{subfigure}%
            \hfill
            \begin{subfigure}[b]{.25\textwidth}
                \centering
                \includegraphics[width=\textwidth]{figures/resolution_test_chart_no_reg_iter_10.png}
                \caption{Without regularizer $N=10$}
                \label{fig:fabioc}
            \end{subfigure}%            
            \hfill
            \begin{subfigure}[b]{.25\textwidth}
                \centering
                \includegraphics[width=\textwidth]{figures/resolution_test_chart_regularizer_008_iter_21.png}
                \caption{With regularizer $N=24$.}
                \label{fig:fabiod}
            \end{subfigure}%            
            \caption{a) has been convolved with a PSF and affected by Poisson shot noise (100 photons expected for the brightest pixel). 
                        Deconvolved b) for 50 iterations without a regularizer c) for 10 iterations without a regularizer d) with Good's roughness regularizer and 
                            converged after 24 iterations with $\lambda=0.03$.}
            \label{fig:image}
        \end{figure}

@@

        Additionally, we notice that the core in the Siemens star is denser is denser 

@@

        In \autoref{fig:rel_energy_regain} we show the relative energy regain $G_\textrm{R}$ \cite{rel_energy_regain} (REG).

@@

        which are further from the ground truth then not reconstructing these frequencies. 

@@

    \subsection{Full Deconvolution}
        In \autoref{tab:deconv} we show the deconvolution time for a 3D dataset with size \todo{size footnote} 
        The basic settings of the different algorithms and packages are noted in \autoref{tab:deconvprop}.
        The aim of this comparison is not to compare the quality of the results but the computing time to achieve similar reconstructions. 
        The algorithms used in \verb|DeconvOptim.jl| are not fundamentally different to a plain Richardson-Lucy deconvolution and therefore the results will 
        not exceed those results significantly in terms of quality.\\
        One should keep in mind that such performance comparisons are usually biased since we are not
        familiar with all options which should be chosen to achieve best performance.
        The comparison should give a rough overview which runtime is expected for each toolbox.\\
        The different packages used were \verb|ThreeDeconv.jl| \cite{ikoma2018convex}, \verb|Huygens| Professional version 21.04 (Scientific Volume Imaging, The Netherlands, \url{http://svi.nl}), \verb|DeconvolutionLab2| \cite{deconvlab2} and \verb|GenericDeconv|.

@@

        For the results of \verb|GenericDeconv| (written in Matlab by Rainer Heintzmann) we used a Poisson loss function with a L-BFGS minimizer 
        and a quadratic non-negativity constraint.
        When the TV regularizer was added the weight was $\lambda = 0.002$.
        The general idea of \verb|GenericDeconv| is very similar to \verb|DeconvOptim.jl|.
        \verb|DeconvolutionLab2| offers a wide variety of different algorithms which however only run on a CPU. 
        To put the numbers into context, even the non-iterative Tikhonov Regularization Inverse Filter (TRIF/TR) had a 
        runtime of \SI{4}{\second}. In the table we show the results of an 
        iterative Richardson-Lucy deconvolution. As number of iterations we set 80 and as regularizer weight $\lambda=0.002$.\\
        In \verb|DeconvOptim.jl| we measured the runtime of the default deconvolution routine with a Poisson loss function
        and the runtime of straightforward Richardson-Lucy deconvolution without a regularizer.
        \verb|ThreeDeconv.jl| is another Julia deconvolution toolbox for deconvolution of three dimensional data.
        In contrast to their according publication \cite{ikoma2018convex} we needed more iterations to achieve good results
        on our used dataset. The regularizer weight was $\lambda=10^{-7}$. 
        The software requires a flat and a dark field image we could not provide. That might influence the quality of the reconstructed images
        but in terms of performance the runtimes might be only by a factor of 2-5 smaller than the runtimes we obtained (Ikoma et al. used 140 iterations in their datasets).

        Since it is hard to judge the quality of those results with a numerical loss function value, we compare them visually.
        The CPU versions were multi threaded (if possible). \verb|DeconvolutionLab2|, \verb|Huygens| and \verb|GenericDeconv| were tested under Windows 10 and not Ubuntu.

        \begin{table}[h]
            \begin{tabular}{l l l}
                & Loss & iterations \\ 
            \hline
            \verb|DeconvOptim.jl| & Poisson & 45 \\
            \verb|DeconvOptim.jl| TV & Poisson & 45\\
            Richardson-Lucy & Poisson & 300 \\
            \verb|ThreeDeconv.jl|  & Gauss+Poisson& 300\\
            \verb|GenericDeconv|  & Poisson & 45\\
            \verb|GenericDeconv| TV & Poisson & 45 \\
            \verb|Huygens| & & \\
            \verb|DeconvolutionLab2| LR & Poisson & 300\\
            \verb|DeconvolutionLab2| LR TV& Poisson & 300 
            \end{tabular}
            \caption{Overview which loss function (noise model) and number of iterations are used.}
            \label{tab:deconvprop}
        \end{table}
        As summary of those performance tests we can state that our Julia implementation is very competitive and especially on the GPU
        it was the fastest of the tested packages. On the CPU it is only beaten by the commercial \verb|Huygens| software. 
        Despite \verb|ThreeDeconv.jl| written in Julia it is notably slower in comparison to our implementation.
        \verb|DeconvolutionLab2| is written in Java and uses hand optimized gradients but still its speed is poor.

>>>>

\begin{abstract}
Deconvolution is a versatile method to enhance the quality of signals measured with systems which can 
be expressed mathematically as a convolution of a system's response function with a signal. 
In this paper, we present \verb|DeconvOptim.jl|, a flexible toolbox written in Julia to deconvolve one or multiple multi-dimensional signals which have been
degraded by a multi-dimensional signal response function. 
\verb|DeconvOptim.jl| works both on CPUs and GPUs and utilizes recent advancements in Julias automatic differentiation ecosystem.

@@

@@

\subsection{Sliding Convolution Windows}
Convolutions can be also calculated with a sliding kernel in real space which is the direct approach to
solve the integral in \autoref{eq:conv}.
That mechanism is the working principle
behind the convolutions in neural nets. The reason is that this operation can be parallelized easily and therefore 
is suited for GPUs. However, the computational complexity for a 2D kernel with size $K \times L$ applied on an image with
$N \times M$ data points is $\mathcal O(N \cdot M \cdot K \cdot L)$. It is clear that for larger kernel sizes ($K,L \gg 1$) the sliding method 
has worse computational complexity than the FFT based convolutions. The many cores of a GPU can mitigate that to a certain extent
but since the FFT (CuFFT in the CUDA library \cite{CUDA}) can be also executed in parallel fashion, both approaches profit from parallelization.
If the kernel is separable over the dimensions (like a Gaussian kernel) the complexity changes to $\mathcal O(N \cdot M \cdot L +  N \cdot M \cdot K)$ however most optical PSFs are not separable and therefore this is not generally applicable.

\subsection{Wrap-Around Artifacts in FFT based Convolutions}
    One drawback of a FFT based convolution are wrap-around artifacts (see arrow in \autoref{fig:withoutpad}). The reason is, that the FFT operation calculates the discrete
    Fourier transform (DFT) which inherently assumes periodic data. If one applies a FFT on a finite data set the assumption is that
    this finite data set repeats itself. 
    To weaken the wrap-around artifacts in the convolution one can zero or mean pad the image before convolution 
    and remove the padding afterwards. Such approaches still suffer from artifacts when boundary values do not fit
    to the mean or to zero as it can be observed in \autoref{fig:withpad} where the arrow is pointing to.
    \begin{figure}[h]
        \centering
        \begin{subfigure}[b]{.25\textwidth}
            \centering
            \begin{tikzpicture}[scale=0.7]
                \node at (0,0) {\includegraphics[width=.9\textwidth]{figures/full_house_no_pad.png}};
                \draw[red] (-2.0, -1.6) rectangle (-1.6, -1.2);
                \node[color=red] at (0,0) {\includegraphics[width=.4\textwidth]{figures/edge_house_no_pad.png}};
                \draw[color=red] (-0.91, -0.9) rectangle (0.9, 0.9); 
                \draw[color=red] (-2.0, -1.2) -- (-0.9, 0.9);
                \draw[color=red] (-1.6, -1.6) -- (0.9, -0.9);
                \draw[<-, color=red, thick ](-0.85, 0) -- (-0.4, 0.6);
            \end{tikzpicture}
            \caption{Without padding}
            \label{fig:withoutpad}
        \end{subfigure}%
        \begin{subfigure}[b]{.25\textwidth}
            \centering
            \begin{tikzpicture}[scale=0.7]
                \node at (0,0) {\includegraphics[width=.9\textwidth]{figures/full_house_pad.png}};
                \draw[red] (-2.0, -1.6) rectangle (-1.6, -1.2);
                \node[color=red] at (0,0) {\includegraphics[width=.4\textwidth]{figures/edge_house_pad.png}};
                \draw[color=red] (-0.90, -0.9) rectangle (0.9, 0.9); 
                \draw[color=red] (-2.0, -1.2) -- (-0.9, 0.9);
                \draw[color=red] (-1.6, -1.6) -- (0.9, -0.9);
                \draw[<-, color=red, thick ](-0.8, -0.7) -- (-0.3, -0.3);
            \end{tikzpicture}
            \caption{With padding.}
            \label{fig:withpad}
        \end{subfigure}
        \caption{Comparison between two blurred images. The image a) is a straightforward FFT based blurred image.
        The zoomed region shows dark artifacts at the left side originated from the right hand side. Also at the top of the image some dark regions from
        the bottom scattered into the bright sky.
        Image b) was padded before blurring with 10 pixels on each side with value 0.7 which results in bright regions around the image. The blur kernel was a Gaussian kernel
        with standard deviation $\sigma=\SI{3}{pixel}$.}
        \label{fig:}
    \end{figure}
    To solve for the wrap-around artifacts in deconvolution we use a different approach which will be presented in \autoref{sec:wrap}. 
\subsection{Preferred Convolution}
The current state of \verb|DeconvOptim.jl| uses the FFT approach to calculate the convolution.
However, in principle one could add the possibility to choose spatially based convolutions.
But since the runtime gets especially long in multi-dimensional signal deconvolution where FFT based methods
are faster, we always use FFT based convolutions here.

\section{Inverse Modelling}
In \autoref{fig:pipeline} we show our approach to deconvolution.
The general idea is to create a (physical) forward model describing the measurement process.
\begin{figure}[h]
    \centering
    \includegraphics[width = .4\textwidth]{figures/pipeline.pdf}
    \caption{General pipeline of our deconvolution approach.}
    \label{fig:pipeline}
\end{figure}

At the beginning of the optimization we start with an initial guess $x(\mathbf r)$. 
Via a mapping function we can impose some restrictions to the final solution (e.g. non-negativity, value intervals).
The forward model is then applied to the mapped reconstruction. We provide a default function where 
the forward function is a convolution but we also allow to plug in a different forward model.
This output of the forward function is compared
to our noisy measurement under a given loss function.
Since deconvolution is an ill-posed problem, we regularize the loss function with a regularizer.
Afterwards we calculate the gradient of the loss function with respect to the reconstruction $x(\mathbf r)$ via automatic differentiation (AD).
The values of the total loss and the gradient are then used by an optimizer (to minimize the total loss) to apply a gradient descent optimization to get a proper reconstruction.\\

As typical for Julia packages, except \verb|FFTW.jl| and \verb|CUDA.jl| \cite{besard2018juliagpu}, \cite{LLVM.jl-2017}, all code is written in Julia.
Therefore it is easy to replace the forward model with any Julia code, as long as we can 
automatically differentiate through the code. The gradient step is currently calculated with the reverse mode AD of \verb|Zygote.jl| \cite{Zygote.jl-2018}. 
As default for the optimization we use the L-BFGS \cite{LBFGS} routine provided by \verb|Optim.jl| \cite{mogensen2018optim}.
L-BFGS is suitable for optimization with many million parameters as it only stores a sparse representation of the inverse Hessian matrix,
which is built from a finite number of past function and gradient evaluations. 
Results are often good after 10 to 50 iterations whereas Richardson-Lucy takes
a few hundreds iterations. 
However, as we will see later due to the more complex optimization routine the speed-up in using L-BFGS is for a simple 
deconvolution minor.

\subsection{Mapping Function}
    Mapping functions are an easy way to constrain the resulting reconstruction to predefined ranges.
    For example, if the optimizer optimizes the variables $\text{rec}$ which are then squared, we know that
    $\text{rec}^2$ is definitely non-negative. Instead of directly using $\text{fwd}(\text{rec})$ we compare
    $\text{fwd}(\text{rec}^2)$ with the measurement $Y$ where $\text{fwd}$ is the function describing the physical operation (mostly 
            a convolution). 
    This mapping function does not change the outcome as long
    as it is an surjective function for the desired number space (e.g. a parabola is non-negative and covers all positive values) 
    and does not change the convexity of a problem.
    Functions like shifted $\tanh$ are possible as well since it is an bijective function in $[-1, 1]$ and monotonic increasing
    in the whole definition space.
    Hence the final result is $\text{rec}^2$ in case of the quadratic mapping.

\subsection{Wrap-around Artifacts Suppression}
    \label{sec:wrap}
    To prevent that wrap-around artifacts occur during deconvolution we can instead
    of reconstructing an object having the same size as the measurement, reconstruct
    an object having a slightly larger size. This is shown in \autoref{fig:wrap_around}.
    \begin{figure}[h]
        \centering
        \begin{tikzpicture}[scale=0.7]
            \draw[fill=red!20!white] (0,0) rectangle (4,4);
            \node[color=red] at (2, 3.8) {Padded area};
            \draw[fill=gray!40!white] (0.5,0.5) rectangle (3.5,3.5); 
            \node[text width=2.0cm, color=black] at (2.1, 2) {area compared with measurement};
        \end{tikzpicture}
        \caption{Reconstructed data is padded before deconvolution but only the red region is compared with the measurement.}
        \label{fig:wrap_around}
    \end{figure}
    Consequently, we only compare the grey part of the reconstruction with the measurement.
    Through the AD and the convolution operations we still optimize for the whole area (including the red one).
    With that padding we reduce that images on the boundary influence in circular manner on reconstruction in the grey area.

\section{Loss Function}
    Apart from forward model, the loss function is the most critical part in the optimization.
    It has to account for the noise created by the measurement process or instrument.

@@

\section{Experiments}
    In this section we show some experiments and compare parts of the software to other solutions.
    All experiments were executed on Ubuntu 20.04 using a AMD Ryzen 5 5600X 6-Core processor (12 threads) with \SI{32}{GB} DDR4 RAM and Julia 1.6.1.
    If possible, all threads were used.
    As GPU accelerator we used a GeForce RTX 2060 SUPER with \SI{8}{GB} memory and CUDA 11.3 and CUDA.jl 3.3.1.
    Our experiments additionally made use of the following packages: ArrayInterface.jl\cite{DifferentialEquations.jl-2017}, 
    BenchmarkTools.jl\cite{BenchmarkTools.jl-2016}, FFTW.jl\cite{FFTW05}, LLVM.jl\cite{LLVM.jl-2017}, Optim.jl\cite{mogensen2018optim}
    and  Zygote.jl\cite{Zygote.jl-2018}. 
    \subsection{One-Dimensional Time Series}
        First, we want to demonstrate the deconvolution of an artificial one dimensional dataset blurred with a certain PSF.
        Additionally, we want to demonstrate the mechanism to reduce wrap-around artifacts in the deconvolution.
        On the left hand side in \autoref{fig:1D_dataset} we can see a data series which has been simulated 
        as a convolution with a PSF and degradation with additive Gaussian noise.
        \begin{figure}[H]
            \begin{subfigure}[b]{.25\textwidth}
                \centering
                \includegraphics[width=\textwidth]{figures/1d_plot_data_blurred.pdf}
                \caption{}
                \label{fig:1D_dataseta}
            \end{subfigure}%
            \begin{subfigure}[b]{.25\textwidth}
                \centering
                \includegraphics[width=\textwidth]{figures/1d_plot_data_res_pad_vs_nopad.pdf}
                \caption{}
                \label{fig:1D_datasetb}
            \end{subfigure}
            \caption{1D data which has been convolved with a PSF and was degraded with additive Gaussian noise with $\sigma=1.0$. Note that for the deconvolution we have chosen a position interval which is well within the full series so that there are not wrap-around artifacts in the measurement.}
            \label{fig:1D_dataset}
        \end{figure}
        On the right hand we can see the results of the following deconvolution:
        \begin{lstlisting}[language=Julia]
r_p, o = deconvolution(measured, psf, loss=Gauss(),
     iterations=12, regularizer=TV(num_dims=1),
     λ=0.05, padding=1)
r, o = deconvolution(measured, psf, loss=Gauss(), 
     iterations=12, regularizer=TV(num_dims=1),
     λ=0.05, padding=0)
        \end{lstlisting}
    The first statement applies a total padding of $1$ meaning that the size is doubled and therefore the measured data is padded.
    As we can see, the blue curve in \autoref{fig:1D_datasetb} does not show these intense spikes at the boundaries as the green curve shows.
    Also we notice that the deconvolved data in \autoref{fig:1D_datasetb} does not show all the details which are visible in the ground truth.
    The reason is that the PSF is a low pass filter and therefore only frequencies which have been passed through the filter 
    can be enhanced in terms of contrast in a deconvolved result.

    \subsection{Influence of Regularizer}
        In this part we want to demonstrate the influence of the regularizer.
        \autoref{fig:fabioa} is a simulated imaged which has been affected by Poisson shot noise and blurring by a PSF.
        \begin{figure}[h]
            \begin{subfigure}[b]{.25\textwidth}
                \centering
                \includegraphics[width=0.85\textwidth]{figures/resolution_test_chart_blurry.png}
                \caption{Measured with Poisson noise}
                \label{fig:fabioa}
            \end{subfigure}%
            \hfill
            \begin{subfigure}[b]{.25\textwidth}
                \centering
                \includegraphics[width=0.85\textwidth]{figures/resolution_test_chart_no_reg_iter_50.png}
                \caption{Without regularizer $N=50$}
                \label{fig:fabiob}
            \end{subfigure}%
            \hfill
            \begin{subfigure}[b]{.25\textwidth}
                \centering
                \includegraphics[width=0.85\textwidth]{figures/resolution_test_chart_no_reg_iter_10.png}
                \caption{Without regularizer $N=10$}
                \label{fig:fabioc}
            \end{subfigure}%            
            \hfill
            \begin{subfigure}[b]{.25\textwidth}
                \centering
                \includegraphics[width=0.85\textwidth]{figures/resolution_test_chart_regularizer_008_iter_21.png}
                \caption{With regularizer $N=24$.}
                \label{fig:fabiod}
            \end{subfigure}%            
            \caption{a) has been convolved with a PSF and affected by Poisson shot noise (100 photons expected for the brightest pixel). 
                        Deconvolved b) for 50 iterations without a regularizer c) for 10 iterations without a regularizer d) with Good's roughness regularizer and 
                            converged after 24 iterations with $\lambda=0.03$.}
            \label{fig:image}
        \end{figure}

@@

        Additionally, we notice that the core in the Siemens star is denser 

@@

        In \autoref{fig:rel_energy_regain} we show the relative energy regain $G_\textrm{R}$ (REG) \cite{rel_energy_regain}.

@@

        which are further from the ground truth than not reconstructing these frequencies. 

@@

    \subsection{Full Deconvolution}
        In \autoref{tab:deconv} we show the deconvolution time for a 3D dataset with a size of $512 \times 256 \times 128$ 
        \footnote{Dataset was taken from \url{http://bigwww.epfl.ch/deconvolution/data/microtubules/}\cite{deconvlab2}.} 
        The basic settings of the different algorithms and packages are noted in \autoref{tab:deconvprop}.
        The aim of this comparison is not to compare the quality of the results but the computing time to achieve similar reconstructions. 
        The algorithms used in \verb|DeconvOptim.jl| are not fundamentally different to a plain Richardson-Lucy deconvolution and therefore the results will 
        not exceed those results significantly in terms of quality.\\
        One should keep in mind that such performance comparisons are usually biased since we are not
        familiar with all options which should be chosen to achieve best performance.
        The comparison should give a rough overview which runtime is expected for each toolbox.\\
        The different packages used were \verb|ThreeDeconv.jl| \cite{ikoma2018convex}, \verb|Huygens| Professional version 21.04 (Scientific Volume Imaging, The Netherlands, \url{http://svi.nl}), \verb|DeconvolutionLab2| \cite{deconvlab2} and \verb|GenericDeconv|.

@@

        For the results of \verb|GenericDeconv| (written in Matlab by Rainer Heintzmann) we used a Poisson loss function with a L-BFGS minimizer 
        and a quadratic non-negativity constraint.
        When the TV regularizer was added the weight was $\lambda = 0.001$.
        The general idea of \verb|GenericDeconv| is very similar to \verb|DeconvOptim.jl|.
        \verb|DeconvolutionLab2| offers a wide variety of different algorithms which however only run on a CPU. 
        In the table we show the results of an 
        iterative Richardson-Lucy deconvolution. The regularizer weight was $\lambda=0.001$.\\
        In \verb|DeconvOptim.jl| we measured the runtime of the default deconvolution routine with a Poisson loss function
        and the runtime of straightforward Richardson-Lucy deconvolution without a regularizer.\\
        \verb|ThreeDeconv.jl| is another Julia deconvolution toolbox for deconvolution of three dimensional data, the regularizer weight was $\lambda=10^{-7}$. 
        Since it is hard to judge the quality of those results with a numerical loss function value, 
        we compare them by the NCC, see \autoref{img:results}.
        The CPU versions were multi threaded (if possible). 
        \verb|DeconvolutionLab2|, \verb|Huygens| and \verb|GenericDeconv| were tested under Windows 10 and not Ubuntu.

        \begin{table}[h]
            \begin{tabular}{l l l}
                & Loss & iterations \\ 
            \hline
            \verb|DeconvOptim.jl| & Poisson & 45 \\
            \verb|DeconvOptim.jl| TV & Poisson & 45\\
            Richardson-Lucy & Poisson & 300 \\
            \verb|ThreeDeconv.jl|  & Gauss+Poisson& 300\\
            \verb|GenericDeconv|  & Poisson & 45\\
            \verb|GenericDeconv| TV & Poisson & 45 \\
            \verb|Huygens| & & \\
            \verb|DeconvolutionLab2| LR & Poisson & 300\\
            \verb|DeconvolutionLab2| LR TV& Poisson & 300 
            \end{tabular}
            \caption{Overview which loss function (noise model) and number of iterations are used.}
            \label{tab:deconvprop}
        \end{table}
        As summary of those performance tests we can state that our Julia implementation is very competitive and especially on the GPU
        it was the fastest of the tested packages. 
        We also see that our routine was faster than a plain Richardson-Lucy mainly due to the much lower number
        of iterations.
        On the CPU \verb|DeconvOptim.jl| is only beaten by the commercial \verb|Huygens| software. 
        Despite \verb|ThreeDeconv.jl| written in Julia it is notably slower in comparison to our implementation.
        \verb|DeconvolutionLab2| is written in Java and uses hand optimized gradients but still its speed is poor in
        comparison to all other packages.
